---
title: "workshop4_pp"
output: html_document
---
```{r}
library(caret)
library(DT)
library(tidyverse)
library(e1071)
library(gridExtra)
library(mda)
library(randomForest)

```

```{r}
install.packages(c("tidyverse", "DT", "gridExtra"))
```

```{r}
install.packages(c("caret", "e1071", "rpart", "randomForest"))
```
```{r}
TBnanostring <- readRDS("TBnanostring.rds")

```

```{r}
View(TBnanostring)
```

```{r}
summary(TBnanostring)
```
```{r}
# too large
plot(TBnanostring)
```

```{r}
library(tidyverse)

pca_out <- prcomp(TBnanostring[, -1])

pca_reduction <- as.data.frame(pca_out$x)
pca_reduction$Condition <- as.factor(TBnanostring$TB_Status)

ggplot(pca_reduction, aes(x = PC1, y = PC2, color = Condition)) +
  geom_point() +
  xlab("PC 1") +
  ylab("PC 2") +
  ggtitle("PCA Plot") +
  theme(plot.title = element_text(hjust = 0.5))
```
# 1. Split the dataset into "training" and "testing" sets using a 70/30 partition (use `set.seed(0)` and the `caret::createDataPartition`).
```{r}
set.seed(0)
test_index <- createDataPartition(TBnanostring$TB_Status, times = 1, 
                                  p = 0.3, list = FALSE)
test_set <- TBnanostring[test_index, ]
train_set <- TBnanostring[-test_index, ]
```

Check data

```{r}
dim(TBnanostring)
dim(train_set)
dim(test_set)

table(TBnanostring$TB_Status)
table(train_set$TB_Status)
table(test_set$TB_Status)
```
Check "LTBI" in the TB_Status column

```{r}
head(TBnanostring[TBnanostring$TB_Status == "LTBI", ])
```

# 2. Apply a Support Vector Machine to these data (try linear, radial, and polynomial kernels).

## Apply SVM with linear kernel
```{r}
svm_linear <- svm(TB_Status ~ .,
                  data = train_set,
                  kernel = "linear",
                  cost = 10)

pred_linear <- predict(svm_linear, test_set)

confusionMatrix(pred_linear, test_set$TB_Status)
```

## Apply SVM (Radial basis kernel)
```{r}
# Train data
TBnanostring$TB_Status <- as.factor(TBnanostring$TB_Status)

fit_radial <- svm(TB_Status ~ ., 
                  data = train_set, 
                  scale = FALSE,
                  kernel = "radial", 
                  cost = 5)

print(fit_radial)
```

```{r}
# Test data
pred_radial <- predict(fit_radial, test_set)

confusionMatrix(pred_radial, test_set$TB_Status)
```

## Apply SVM(polynomial kernel)
```{r}

# Train data
TBnanostring$TB_Status <- as.factor(TBnanostring$TB_Status)

fit_polynomial <- svm(TB_Status ~ ., 
                  data = train_set, 
                  scale = FALSE,
                  kernel = "polynomial", 
                  cost = 5)

print(fit_polynomial)
```


```{r}
# Test data
pred_polynomial <- predict(fit_polynomial, test_set)

confusionMatrix(pred_polynomial, test_set$TB_Status)
```


# Boxplot
```{r}
library(tidyverse)

TBnanostring %>%
  pivot_longer(
    cols = -TB_Status,
    names_to = "Gene",
    values_to = "Expression"
  ) %>%
  ggplot(aes(x = TB_Status, y = Expression, fill = TB_Status)) +
  geom_boxplot() +
  facet_wrap(~ Gene, scales = "free", ncol = 4) +
  theme(axis.text.x = element_blank(),
        legend.position = "bottom") +
  xlab("TB Status") +
  ylab("Gene Expression") +
  ggtitle("Gene Expression by TB Status")
```
```{r}


# Get all gene column numbers, excluding TB_Status
gene_cols <- 2:ncol(TBnanostring)

# Loop every 9 genes
for (start_col in seq(2, ncol(TBnanostring), by = 9)) {
  
  end_col <- min(start_col + 8, ncol(TBnanostring))
  
  p <- TBnanostring %>%
    select(TB_Status, start_col:end_col) %>%
    pivot_longer(
      cols = -TB_Status,
      names_to = "Gene",
      values_to = "Expression"
    ) %>%
    ggplot(aes(x = TB_Status, y = Expression, fill = TB_Status)) +
    geom_boxplot() +
    facet_wrap(~ Gene, scales = "free", ncol = 3) +
    theme(axis.text.x = element_blank(),
          legend.position = "bottom") +
    xlab("TB Status") +
    ylab("Gene Expression") +
    ggtitle(paste("Gene Expression by TB Status: Columns", start_col, "to", end_col))
  
  print(p)
}
```

#3. Apply a Random Forest Model to these data.

## Trainning
```{r}
TB_classifier <- randomForest(TB_Status~., data = train_set, 
                    importance=T)
TB_classifier
```

## Testing
```{r}
predicted_TB_table <- predict(TB_classifier, test_set)
confusionMatrix(predicted_TB_table, test_set$TB_Status)

```

5. Compare the overall accuracy of the prediction methods for each of the machine learning tools in the previous problem. Which one performs the best?  

(Note: the `TBnanostring.Rmd` and `TBnanostring.html` files provide suggested solutions for these analyses)

## Summary
```{r}
svm_results <- data.frame(
  Model = c("SVM Linear", "SVM Radial", "SVM Polynomial", "Random Forest"),
  Accuracy = c(0.9259, 0.9074, 0.8889, 0.8519),
  Sensitivity = c(0.9167, 0.9167, 0.7917, 0.8750),
  Specificity = c(0.9333, 0.9000, 0.9667, 0.8333),
  Balanced_Accuracy = c(0.9250, 0.9083, 0.8792, 0.8542),
  Kappa = c(0.8500, 0.8133, 0.7712, 0.7025),
  Correct_Total = c("50 / 54", "49 / 54", "48 / 54", "46 / 54")
)

svm_results
```